---
title: "01_stats"
format: html
editor: visual
editor_options: 
  chunk_output_type: console
---

## Packages

```{r}
library(dplyr)
library(ggplot2)
library(readxl)
library(purrr)
library(tidyr)
library(forcats)
library(ggeasy)
library(cowplot)

change_colname <- function(df, a, b){
  i <- grep(a, names(df))[1]
  names(df)[i] <- b
  df
}


```

## Data

```{r}

#
# a Main data (polygons for "deposition" i.e. built-out areas after 2004)
#

# reading excel causes encoding trouble!
# fn <- "data/kystnatur_statistics.xlsx"
# readxl::excel_sheets(fn)
# dat_all <- read_excel(fn)
# a <- grep("^Havne", dat_all$objtype)
# b <- a[1]
# dat_all$objtype[b]
# dat_all[b,]
# dat_all[b,1]

fn <- "data/kystnatur_statistics_deposition.csv"
dat_all <- readr::read_csv2(fn)


fn <- "data/kystnatur_statistics.xlsx"

#
# b Kommune (area, no. of inhabitats) ----
#
# kommune: 
lookup_komm <- read_excel(fn, sheet = "kommuner")

lookup_komm <- change_colname(lookup_komm, "Befolkning per 1.1. (personer)", "befolkning")
lookup_komm <- change_colname(lookup_komm, "Areal (km2)", "kommuneareal")

#
# c Hytter (cottages) ----
#
# 03174: Eksisterende bygningsmasse. Fritidsbygg, etter region, år og bygningstype
# https://www.ssb.no/statbank/table/03174
fn <- "data/ssb/03174_20260410-144751.xlsx"
lookup_cabins <- read_excel(
  fn, skip = 5, n_max = 357, 
  col_names = c("kommune", "hytter", "fritidsbolig")) %>%
  mutate(
    kommunenum = substr(kommune, 3, 6),
    hytter_og_fritidsboliger = hytter + fritidsbolig) 
# lookup_cabins %>%  head()

#
# . d Næring ----
#
# 07984: Sysselsatte per 4. kvartal, etter region, næring (SN2007), år og alder
# Oppdatert: 04.02.2026, 08:00
# https://www.ssb.no/statbank/table/03174
# also see: https://www.ssb.no/klass/sok?query=kommune&includeCodelists=false&ssbSection= 
fn <- "data/ssb/07984_20260410-133349.xlsx"
read_excel(fn, skip = 3)
# lookup_work_long <- read_excel(fn, skip = 3, n_max = 6480, col_names = c("kommune", "næring", "sysselsatte_char"))

lookup_work_long <- read_excel(fn, skip = 3, n_max = 6480)
names(lookup_work_long) <- c("kommune", "naering", "sysselsatte_char")
lookup_work_long

# fill empty 'kommune' fields
for (i in 2:nrow(lookup_work_long)){
  if (is.na(lookup_work_long$kommune[i]))
    lookup_work_long$kommune[i] <- lookup_work_long$kommune[i-1]
  }

lookup_work_long <- lookup_work_long %>% 
  filter(!is.na(naering)) %>% 
  mutate(
    kommunenum = substr(kommune, 3, 6),
    sysselsatte = as.numeric(sysselsatte_char),
    sysselsatte = ifelse(is.na(sysselsatte), 0, sysselsatte)
  )

lookup_work_wide <- lookup_work_long %>% 
  select(kommunenum, naering, sysselsatte) %>% 
  pivot_wider(names_from = naering, values_from = sysselsatte)

# add sum
# lookup_work_wide$n_sysselsatte <- apply(lookup_work_wide[,-(1:2)], 1, sum)

names(lookup_work_wide)
lookup_work_wide <- change_colname(lookup_work_wide, "00-99 Alle n", "n_alle")
lookup_work_wide <- change_colname(lookup_work_wide, "01-03 Jordbruk, skogbruk og fiske", "n_primaernaering")
lookup_work_wide <- change_colname(lookup_work_wide, "05-09 Bergverksdrift og utvinning", "n_bergverk")
lookup_work_wide <- change_colname(lookup_work_wide, "10-33 Industri", "n_industri")

lookup_work <- lookup_work_wide |> 
  mutate(
    n_industri = n_bergverk + n_industri,
    prosent_primaernaering = n_primaernaering/n_alle*100 %>% round(2),
    prosent_industri = n_industri/n_alle*100 %>% round(2)
  ) %>%
  select(kommunenum, n_primaernaering, n_industri, n_alle, prosent_primaernaering, prosent_industri)

ggplot(lookup_work, aes(prosent_primaernaering, prosent_industri)) +
  geom_point()

```

## Types  

```{r}

xtabs(~objtype, dat_all) 

gg1 <- dat_all %>% 
  count(objtype) %>% 
  arrange(desc(n)) %>%
  mutate(objtype = fct_inorder(objtype)) %>% 
  ggplot(aes(objtype, n)) +
  geom_col() +
  easy_rotate_x_labels(-45) +
  labs(main = "Type by number")

gg2 <- dat_all %>% 
  count(objtype, wt = area, name = "area") %>% 
  arrange(desc(area)) %>%
  mutate(
    objtype = fct_inorder(objtype),
    area = area/1E3) %>% 
  ggplot(aes(objtype, area)) +
  geom_col() +
  easy_rotate_x_labels(-45) +
  labs(main = "Type by area")

# gg2

plot_grid(gg1, gg2, ncol = 1)

```

## Area and number stats  

### Total data set  
```{r}

cat("Number of structures: ")
nrow(dat_all)

cat("Combined area of structures (km2): ")
sum(dat_all$area)/1000000

```

### Data excluding bridges and "harbour area polygons"  
```{r}

dat_all2 <- dat_all %>% 
  filter(!(grepl("Bru",objtype) | grepl("Havneomr",objtype)))

cat("Number of structures: ")
nrow(dat_all2)

cat("Combined area of structures (km2): ")
sum(dat_all2$area)/1000000

```

## Also exclude floating piers (flytebrygge)  
Basis for the rest of the statistics by kommune etc.  

### Filter and sort
```{r}

# Filter and sort 
dat_sort <- dat_all %>% 
  filter(!(grepl("Bru",objtype) | grepl("Havneomr",objtype) | grepl("Flytebrygge",objtype))) %>% 
  arrange(desc(area)) %>% 
  mutate(cumularea = cumsum(area)/1000)

# # Show and remove "Havneområde" in Sandefjord sel <- dat_sort$ID == 199033
# dat_sort[sel,] dat_sort <- dat_sort[!sel,]
#
# # Remove "Havneområde" in Hammerfest #
# https://www.norgeibilder.no/?x=819474.6622&y=7863484.4902&level=12&utm=33&projects=&layers=&plannedOmlop=0&plannedGeovekst=0
# sel <- dat_sort$ID == 198637 dat_sort[sel,] dat_sort <- dat_sort[!sel,]
#
# # Remove "Havneområde" in Snefjord, Måsøy #
# https://www.norgeibilder.no/?x=850190.8275&y=7881841.4665&level=12&utm=33&projects=&layers=&plannedOmlop=0&plannedGeovekst=0
# sel <- dat_sort$ID == 198664 dat_sort[sel,] dat_sort <- dat_sort[!sel,]
#
# # Remove "Havneområde" in Havøysund, Måsøy #
# https://www.norgeibilder.no/?x=849936.908&y=7904799.4875&level=12&utm=33&projects=&layers=&plannedOmlop=0&plannedGeovekst=0
# sel <- dat_sort$ID == 198659 dat_sort[sel,] dat_sort <- dat_sort[!sel,]
#
# # Remove "Havneområde" in Havøysund, Måsøy #
# https://www.norgeibilder.no/?x=1040129.7057&y=7902224.3643&level=12&utm=33&projects=&layers=&plannedOmlop=0&plannedGeovekst=0
# sel <- dat_sort$ID == 198641 dat_sort[sel,] dat_sort <- dat_sort[!sel,]

```

### Final stats, excluding floating piers   
```{r}

cat("Number of structures: ")
nrow(dat_sort)

cat("Combined area of structures (km2): ")
sum(dat_sort$area)/1000000

```



### Floating piers (flytebrygge)  
```{r}

dat_all2 <- dat_all %>% 
  filter(grepl("Flytebrygge",objtype))

cat("Number of floating piers: ")
nrow(dat_all2)

cat("Combined area of floating piers (km2): ")
sum(dat_all2$area)/1000000

rectangle_length <- dat_all2 %>% 
  mutate(
    side_a = side_a(area, length),
    side_b = side_b(area, length)
  ) |> 
  pull(
    side_a)

cat("Combined length of floating piers (km): ")
sum(rectangle_length, na.rm = TRUE)/1000

cat("Median length of floating piers (m): ")
median(rectangle_length, na.rm = TRUE)
cat("Mean length of floating piers (m): ")
mean(rectangle_length, na.rm = TRUE)




```


### Types  

```{r}

xtabs(~objtype, dat_sort) 

gg1 <- dat_sort %>% 
  count(objtype) %>% 
  arrange(desc(n)) %>%
  mutate(objtype = fct_inorder(objtype)) %>% 
  ggplot(aes(objtype, n)) +
  geom_col() +
  easy_rotate_x_labels(-45) +
  labs(main = "Type by number")

gg2 <- dat_sort %>% 
  count(objtype, wt = area, name = "area") %>% 
  arrange(desc(area)) %>%
  mutate(
    objtype = fct_inorder(objtype),
    area = area/1E3) %>% 
  ggplot(aes(objtype, area)) +
  geom_col() +
  easy_rotate_x_labels(-45) +
  labs(main = "Type by area")

# gg2

plot_grid(gg1, gg2, ncol = 1)

```

## Combined length all structures

### Length per structure   
Assuming structures are rectangle-shaped  
```{r}

# Example 1
#  _ _ _ _ _ _ 
# |           |
# |           |
# |_ _ _ _ _ _|
# perimeter = 6*2 + 3*2 = 18
# area = 6*3 = 18
#
# Example 2
#  _ _ _ _ 
# |        |
# |        |_ _
# |_ _ _ _  _ _|
# perimeter (clockwise from top) = 4 + 2 + 2 + 1 + 6 + 3 = 18
# area = 4*3 + 2*1 = 14
#


# areal A = a*b
# omkrets P = 2*a + 2*b
#         = 2*a + 2*(areal/a)
#         = 2*a + 2*areal/a
#         = 2*(a + areal/a)
# a = (P/2 + sqrt((P/2)^2 - 4*A))/2
side_a <- function(A,P) (P/2 + sqrt((P/2)^2 - 4*A))/2
side_b <- function(A,P) (P/2 - sqrt((P/2)^2 - 4*A))/2
# assume dimension 2 x 4, ie area = 8 and perimeter = 12
side_a(8,12)
side_b(8,12)
# assume dimension 1 x 8, ie area = 8 and perimeter = 18
side_a(8,18)
side_b(8,18)
# example 1 above
side_a(18,18) # 6
side_b(18,18) # 3
# example 2 above
side_a(14,18) # 7
side_b(14,18) # 2
# while real length is 6
# conclusion: concave parts causes real length to be shorter than "rectangle-based" length

dat_sort <- dat_sort |> 
  mutate(
    side_a = side_a(area, length),
    side_b = side_b(area, length)
  )

```

### Combined length and mean width of all structures
```{r}

cat("Combined length of structures (km): ")
sum(dat_sort$side_a, na.rm = TRUE)/1000
# 2340 km

cat("Mean width of structures (m): ")
mean(dat_sort$side_a, na.rm = TRUE)
# 14.7 m

```

## Combined length, Hammerfest sample 

### Check using Hammerfest    

```{r}

library(sf)
check <- sf::read_sf("data/hammerfest_sjekk.gpkg")
check
plot(check["objtype"])
# the whole built out area is about 5-600 meters long

check$id <- 1:nrow(check) 
plot(check["id"])

# Order from largest to smallest
order <- order(check$Shape_Area) |> rev()
order

# Plot the 12 biggest objects
# We need to use good old base plot() and lines()
hf_xlim <- st_bbox(check)[c(1, 3)]
hf_ylim <- st_bbox(check)[c(2, 4)]
order <- order(check$Shape_Area) |> rev()
coords <- st_coordinates(check)
par(mfrow = c(3,4), mar = c(0,0,0,0))
for (i in 1:12){
  # plot(check[order[i], "objtype"], xlim = hf_xlim, ylim = hf_ylim)
  plot(0,0, type = "n", xlim = hf_xlim, ylim = hf_ylim)
  coo <- coords[coords[,"L3"]==order[i],]
  lines(coo[,"X"], coo[,"Y"], )
}

# Cumulative area 
df <- check |> sf::st_drop_geometry()
cumsum <- df |> 
  arrange(desc(Shape_Area)) |> 
  pull(Shape_Area) |> 
  cumsum()

# Cumulative area in percent of total
# 5 largest = 98.8 %
# WHich is reassuring
cumsum/sum(df$Shape_Area)*100
# [1]  84.46944  90.65949  94.74987  96.70993  98.35111  98.81937  99.20412  99.55998  99.77858 

# Calculate ideal length of sides
check$side_a <- side_a(check$Shape_Area, check$Shape_Leng)
check$side_b <- side_b(check$Shape_Area, check$Shape_Leng)

# Check sum
sum(check$side_a)
# 2289 m - seems very high, should be more like 5-600 m

sum(check$side_a[check$side_b >= 2])
# 1493
sum(check$side_a[check$side_b >= 3])
# 1493
sum(check$side_a[check$side_b >= 5])
# 1493
sum(check$side_a[check$side_b >= 10])
# 1244

```

### Merging touching polygons and calculate diameter  
```{r}

# Find polygons sharing boundaries
nb <- st_touches(check)

# Connected components
library(igraph)

g <- graph_from_adj_list(nb)
check$cluster <- components(g)$membership

# Check clusters
plot(check["cluster"])

# Dissolve by cluster
# may need to do this: install.packages("lwgeom")
check2 <- check |>
  dplyr::group_by(cluster) |>
  dplyr::summarise() |> 
  mutate(
    Shape_Area = st_area(geom),
    Shape_Leng = st_perimeter(geom)
  )

# Check clusters
plot(check2)

# Check 
check2$side_a <- side_a(check2$Shape_Area, check2$Shape_Leng)
check2$side_b <- side_b(check2$Shape_Area, check2$Shape_Leng)

sum(check2$side_a)
# 1107 m - much better

sum(check2$side_a[as.numeric(check2$Shape_Area) >= 50])
# 1107 m - much better

library(sf)

# find diameter
get_diameter <- function(poly) {
  hull <- st_convex_hull(poly)
  coords <- st_coordinates(hull)[, 1:2]
  # Remove duplicated last point
  coords <- coords[!duplicated(coords), ]
  max(dist(coords))
}

check2$diameter <- sapply(st_geometry(check2), get_diameter)

sum(check2$diameter)
# 760 m - even better

sum(check2$diameter[as.numeric(check2$Shape_Area) >= 50])
# 621 m

```

## Combined length, all municipalities   
* Estimate using 20% of municipalities   
* Ca. 5 minutes for 20% of municipalities  

```{r}
# Connected components
library(igraph)

library(sf)
sf_all <- sf::st_read("data/build_Construction_clip/build_Construction_clip.shp")
table(sf_all$kommunenum) |> length()
# 225 municipalities

# Pick n municipalities at random:
n <- 44   # 20 % of total number of municipalities

kommunenum_sel <- table(sf_all$kommunenum) |> names() |> sample(n)

sf <- sf_all |> 
  # filter(kommunenav == "Hammerfest") |> 
  filter(kommunenum %in% kommunenum_sel) |> 
  filter(!(grepl("Bru",objtype) | grepl("Havneomr",objtype) | grepl("Flytebrygge",objtype)))

# plot(sf)

# Find polygons sharing boundaries
nb <- st_touches(sf)

# Make 'cluster' for polygons sharing boundaries  
g <- graph_from_adj_list(nb)
sf$cluster <- components(g)$membership

# The next step is a slow one
# We use alternative 1, it seems faster

# ALT 1. Dissolve by cluster using summarise()

# may need to do this: install.packages("lwgeom")

t0 <- Sys.time()
sf2 <- sf |>
  dplyr::group_by(cluster) |>
  dplyr::summarise()
t1 <- Sys.time()
t1-t0  # for Hammerfest: 2.6 seconds

# if polygons share borders but do not overlap (a coverage), is_coverage = true can
# improve speed a lot:
#  dplyr::summarise(is_coverage = true)
# but they DO opverlap

# ALT 2. Use st_union where only the geometric union is executed
# turns out to be slower, see resukt for Hammerfest

# t0 <- Sys.time()
# clusters <- split(sf, sf$cluster)
# sf2 <- do.call(
#   rbind,
#   lapply(names(clusters), function(cl) {
#     st_sf(
#       cluster = cl,
#       geometry = st_union(clusters[[cl]])
#     )
#   })
# )
# t1 <- Sys.time()
# t1-t0  # for Hammerfest: 4.6 seconds

# ALT 2 -end

sf2 <- sf2 |> 
  mutate(
    Shape_Area = st_area(geometry),
    Shape_Leng = st_perimeter(geometry))
    
sf2$side_a <- side_a(sf2$Shape_Area, sf2$Shape_Leng) |> as.numeric()
sf2$side_b <- side_b(sf2$Shape_Area, sf2$Shape_Leng) |> as.numeric()

# sf clusters
# plot(sf2)

sf2$diameter <- sapply(st_geometry(sf2), get_diameter)

```

### Some plots 
```{r}


ggplot(sf2, aes(diameter)) +
  geom_histogram()

ggplot(sf2, aes(side_a)) +
  geom_histogram()

ggplot(sf2 |> filter(as.numeric(Shape_Area) >= 10 & side_a < 5000), aes(diameter/side_a)) +
  geom_histogram()

```

### Total length for sample and for Norway  
```{r}

cat("Total length in km (sample): ",
    sum(sf2$diameter)/1000, "\n")
# 555.8 for 20% av kommunene

# Total length in km, only polygons > 50 m2
cat("Total length in km, only polygons > 50 m2 (sample): ",
    sum(sf2$diameter[as.numeric(sf2$Shape_Area) >= 50])/1000,
    "\n")
# 324.9 for 20%

cat("Total length in km (Norway): ",
    sum(sf2$diameter)/1000*5, "\n")
# [1] 2779

# Total length in km, only polygons > 50 m2
cat("Total length in km, only polygons > 50 m2 (Norway): ",
    sum(sf2$diameter[as.numeric(sf2$Shape_Area) >= 50])/1000*5,
    "\n")
# [1] 1624.5

```

## Combined length, flytebrygger   
* Estimate using 40% of municipalities (ca. 2 minutes)

```{r}
# Connected components
library(igraph)

library(sf)
sf_all <- sf::st_read("data/build_Construction_clip/build_Construction_clip.shp")
table(sf_all$kommunenum) |> length()
# 225 municipalities

# Pick n municipalities at random:
n <- 88   # 20 % of total number of municipalities

kommunenum_sel <- table(sf_all$kommunenum) |> names() |> sample(n)

sf <- sf_all |> 
  # filter(kommunenav == "Hammerfest") |> 
  filter(kommunenum %in% kommunenum_sel) |> 
  filter(grepl("Flytebrygge",objtype))

# plot(sf)

# Find polygons sharing boundaries
nb <- st_touches(sf)

# Make 'cluster' for polygons sharing boundaries  
g <- graph_from_adj_list(nb)
sf$cluster <- components(g)$membership

# The next step is a slow one
# We use alternative 1, it seems faster

# ALT 1. Dissolve by cluster using summarise()

# may need to do this: install.packages("lwgeom")

t0 <- Sys.time()
sf2 <- sf |>
  dplyr::group_by(cluster) |>
  dplyr::summarise()
t1 <- Sys.time()
t1-t0  # for Hammerfest: 2.6 seconds

# if polygons share borders but do not overlap (a coverage), is_coverage = true can
# improve speed a lot:
#  dplyr::summarise(is_coverage = true)
# but they DO opverlap

# ALT 2. Use st_union where only the geometric union is executed
# turns out to be slower, see resukt for Hammerfest

# t0 <- Sys.time()
# clusters <- split(sf, sf$cluster)
# sf2 <- do.call(
#   rbind,
#   lapply(names(clusters), function(cl) {
#     st_sf(
#       cluster = cl,
#       geometry = st_union(clusters[[cl]])
#     )
#   })
# )
# t1 <- Sys.time()
# t1-t0  # for Hammerfest: 4.6 seconds

# ALT 2 -end

sf2 <- sf2 |> 
  mutate(
    Shape_Area = st_area(geometry),
    Shape_Leng = st_perimeter(geometry))
    
sf2$side_a <- side_a(sf2$Shape_Area, sf2$Shape_Leng) |> as.numeric()
sf2$side_b <- side_b(sf2$Shape_Area, sf2$Shape_Leng) |> as.numeric()

# sf clusters
# plot(sf2)

sf2$diameter <- sapply(st_geometry(sf2), get_diameter)

cat("Total length of floating piers in km (Norway): ",
    sum(sf2$diameter)/1000*2.5, "\n")
# [1] 577

```

## Area distribution 

### Quantiles  
```{r}
quantile(dat_sort$area, probs = c(0, 0.2, 0.5, 0.8, 1)) %>% round(1)
quantile(dat_sort$cumularea, probs = c(0, 0.2, 0.5, 0.8, 1))
```

### Fraction of total area is coverd by areas > x  
```{r}
fractions <- c(0, 0.20, 0.40, 0.60, 0.80,1.00)
purrr::map_dfr(
  fractions,
  \(fraq){
    tot <- tail(dat_sort$cumularea,1)*fraq
  dat_sel <- dat_sort %>% 
    filter(cumularea < tot)  
  tibble(
    fraction_of_total_area = fraq,
    n_polygons = nrow(dat_sel),
    perc_polygons = n_polygons/nrow(dat_sort)*100,
    areas_larger_than = tail(dat_sel$area, 1))
  })
  
```


## Stats by kommune  

### Some stats for kommune features  
```{r}

ggplot(lookup_komm, aes(befolkning)) +
  geom_histogram()
lookup_komm %>% 
  mutate(
    storkommune = case_when(
      befolkning >= 20000 ~ "Over 20000",
      befolkning < 20000 ~ "Under 20000")) %>% 
  count(storkommune)

ggplot(lookup_cabins, aes(hytter_og_fritidsboliger)) +
  geom_histogram()

ggplot(lookup_work, aes(prosent_primærnaering)) +
  geom_histogram(binwidth = 0.5)

ggplot(lookup_work, aes(prosent_industri)) +
  geom_histogram(binwidth = 1)

```

### Make data for kommune level
```{r}

threshold_hytte <- 2500

dat_k <- dat_sort %>% 
  group_by(
    kommunenum) %>% 
  summarise(
    antall = n(),
    areal_dekar = sum(area)/1000) %>% 
  left_join(
    lookup_komm) %>% 
  left_join(
    lookup_cabins) %>% 
  left_join(
    lookup_work) %>% 
  arrange(
    desc(areal_dekar)) %>% 
  mutate(
    kommunenavn = fct_inorder(kommunenavn),
    storkommune = case_when(
      befolkning >= 20000 ~ "Over 20000",
      befolkning < 20000 ~ "Under 20000"),
    hytter_og_fritidsboliger_perinnb = hytter_og_fritidsboliger/befolkning,
    hyttekommune = case_when(
      hytter_og_fritidsboliger >= threshold_hytte ~ paste("Over", threshold_hytte, "hytter/fritidsbolig"),
      hytter_og_fritidsboliger < threshold_hytte ~ paste("Under", threshold_hytte, "hytter/fritidsbolig")),
    primaernaeringskommune = case_when(
      prosent_primaernaering >= 10 ~ "Over 10% primaernaering",
      prosent_primaernaering < 10 ~ "Under 10% primaernaering"),
    industrikommune = case_when(
      prosent_industri >= 10 ~ "Over 10% industri/anlegg",
      prosent_industri < 10 ~ "Under 10% industri/anlegg")
  )

ggplot(dat_k[1:50,], aes(kommunenavn, areal_dekar)) +
  geom_col(aes(fill = storkommune)) +
  theme_bw() +
  easy_rotate_x_labels(angle = -45)

```


### Type of kommune - tables  
```{r}

dat_k %>% 
  summarize(
    `Antall kommuner` = n(),
    `Antall inngrep` = sum(antall),
    `Dekar inngrep` = sum(areal_dekar),
    `Dekar per kommune` = `Dekar inngrep`/`Antall kommuner`)

dat_k %>% 
  group_by(storkommune) %>%
  summarize(
    `Antall kommuner` = n(),
    `Antall inngrep` = sum(antall),
    `Dekar inngrep` = sum(areal_dekar),
    `Dekar per kommune` = `Dekar inngrep`/`Antall kommuner`)

dat_k %>% 
  group_by(primaernaeringskommune) %>%
  summarize(
    `Antall kommuner` = n(),
    `Antall inngrep` = sum(antall),
    `Dekar inngrep` = sum(areal_dekar),
    `Dekar per kommune` = `Dekar inngrep`/`Antall kommuner`)

dat_k %>% 
  group_by(industrikommune) %>%
  summarize(
    `Antall kommuner` = n(),
    `Antall inngrep` = sum(antall),
    `Dekar inngrep` = sum(areal_dekar),
    `Dekar per kommune` = `Dekar inngrep`/`Antall kommuner`)

dat_k %>% 
  group_by(hyttekommune) %>%
  summarize(
    `Antall kommuner` = n(),
    `Antall inngrep` = sum(antall),
    `Dekar inngrep` = sum(areal_dekar),
    `Dekar per kommune` = `Dekar inngrep`/`Antall kommuner`)


```

### Type of kommune - combination tables
```{r}

dat_k %>% 
  group_by(storkommune,industrikommune) %>%
  summarize(
    `Antall kommuner` = n(),
    # `Antall inngrep` = sum(antall),
    `Dekar inngrep` = sum(areal_dekar),
    `Dekar per kommune` = `Dekar inngrep`/`Antall kommuner`)

dat_k %>% 
  group_by(storkommune,industrikommune, hyttekommune) %>%
  summarize(
    `Antall kommuner` = n(),
    # `Antall inngrep` = sum(antall),
    `Dekar inngrep` = sum(areal_dekar),
    `Dekar per kommune` = `Dekar inngrep`/`Antall kommuner`)

dat_k %>% 
  group_by(storkommune,industrikommune, hyttekommune, primaernaeringskommune) %>%
  summarize(
    `Antall kommuner` = n(),
    # `Antall inngrep` = sum(antall),
    `Dekar inngrep` = sum(areal_dekar),
    `Dekar per kommune` = `Dekar inngrep`/`Antall kommuner`)

dat_k %>% 
  group_by(industrikommune, primaernaeringskommune) %>%
  summarize(
    `Antall kommuner` = n(),
    # `Antall inngrep` = sum(antall),
    `Dekar inngrep` = sum(areal_dekar),
    `Dekar per kommune` = `Dekar inngrep`/`Antall kommuner`)

```

## Maps  

### Geodata  
```{r}

library(sf)
library(leaflet)
sf <- read_sf("data/deposition/deposition.shp")

# Extract coordinates
coord_utm <- st_coordinates(sf)
sf_dat <- st_drop_geometry(sf) %>% 
  bind_cols(as.data.frame(coord_utm))

sf_datk <- sf_dat %>% 
  filter(!objtype %in% c("Flytebrygge", "Bru", "Havneområde")) %>% 
  group_by(fylkesnavn, kommunenum) %>% 
  summarize(
    area_dekar = sum(area)/1000,
    X = weighted.mean(X, w=area),
    Y = weighted.mean(Y, w=area))  %>% 
  left_join(
    lookup_komm) %>% 
  left_join(
    lookup_cabins) %>% 
  left_join(
    lookup_work) %>% 
  mutate(
    circle_size = sqrt(area_dekar/2))

stem(sf_datk$circle_size)

coord_latlon <- st_as_sf(sf_datk,
                         coords = c("X", "Y"),
                  	     crs = 6173) %>% 
  st_transform(4326) %>% 
  st_coordinates() %>% 
  as.data.frame() %>% 
  set_names(c("lng", "lat"))

sf_datk <- sf_datk %>% 
  bind_cols(coord_latlon)

```

### Leaflet  
```{r}

leaflet(sf_datk %>% filter(befolkning < 20000)) %>%
  # leaflet::addTiles() %>%
  addProviderTiles(providers$Stadia.AlidadeSatellite) %>% 
  addCircleMarkers(lng = ~lng, lat =  ~lat, radius = ~circle_size,
                   color = "#99ccff", fillColor = "#99ccff",
                   popup = ~paste0(kommunenavn, 
                                   "<br>befolkning: ", befolkning)) %>% 
  addCircleMarkers(lng = ~lng, lat =  ~lat, radius = ~circle_size,
                   popup = ~paste0(kommunenavn, 
                                   "<br>befolkning: ", befolkning), 
                   color = "#ff1a1a", fillColor = "#ff1a1a", opacity = 0.8,
                   data = sf_datk %>% filter(befolkning >= 20000))

leaflet(sf_datk %>% filter(prosent_industri < 10)) %>%
  # leaflet::addTiles() %>%
  addProviderTiles(providers$Stadia.AlidadeSatellite) %>% 
  addCircleMarkers(lng = ~lng, lat =  ~lat, radius = ~circle_size,
                   color = "#99ccff", fillColor = "#99ccff",
                   popup = ~paste0(kommunenavn, 
                                   "<br>prosent_industri: ", round(prosent_industri,1))) %>% 
  addCircleMarkers(lng = ~lng, lat =  ~lat, radius = ~circle_size,
                   popup = ~paste0(kommunenavn, 
                                   "<br>prosent_industri: ", round(prosent_industri,1)), 
                   color = "#ff1a1a", fillColor = "#ff1a1a", opacity = 0.8,
                   data = sf_datk %>% filter(prosent_industri >= 10))

treshhold <- 2500
leaflet(sf_datk %>% filter(hytter_og_fritidsboliger < treshhold)) %>%
  # leaflet::addTiles() %>%
  addProviderTiles(providers$Stadia.AlidadeSatellite) %>% 
  addCircleMarkers(lng = ~lng, lat =  ~lat, radius = ~circle_size,
                   color = "#99ccff", fillColor = "#99ccff",
                   popup = ~paste0(kommunenavn,
                                   "<br>hytter_og_fritidsboliger: ", hytter_og_fritidsboliger)) %>%
  addCircleMarkers(lng = ~lng, lat =  ~lat, radius = ~circle_size,
                   popup = ~paste0(kommunenavn,
                                   "<br>hytter_og_fritidsboliger: ", hytter_og_fritidsboliger),
                   color = "#ff1a1a", fillColor = "#ff1a1a", opacity = 0.8,
                   data = sf_datk %>% filter(hytter_og_fritidsboliger >= treshhold))


leaflet(sf_datk %>% filter(prosent_primaernaering < 10)) %>%
  # leaflet::addTiles() %>%
  addProviderTiles(providers$Stadia.AlidadeSatellite) %>% 
  addCircleMarkers(lng = ~lng, lat =  ~lat, radius = ~circle_size,
                   color = "#99ccff", fillColor = "#99ccff",
                   popup = ~paste0(kommunenavn, 
                                   "<br>prosent_primaernaering: ", round(prosent_primaernaering,1))) %>% 
  addCircleMarkers(lng = ~lng, lat =  ~lat, radius = ~circle_size,
                   popup = ~paste0(kommunenavn, 
                                   "<br>prosent_primaernaering: ", round(prosent_primaernaering,1)), 
                   color = "#ff1a1a", fillColor = "#ff1a1a", opacity = 0.8,
                   data = sf_datk %>% filter(prosent_primaernaering >= 10))


```


```{r}

# 15 instead of 10
leaflet(sf_datk %>% filter(prosent_industri < 15)) %>%
  # leaflet::addTiles() %>%
  addProviderTiles(providers$Stadia.AlidadeSatellite) %>% 
  addCircleMarkers(lng = ~lng, lat =  ~lat, radius = ~circle_size,
                   color = "#99ccff", fillColor = "#99ccff",
                   popup = ~paste0(kommunenavn, 
                                   "<br>prosent_industri: ", round(prosent_industri,1))) %>% 
  addCircleMarkers(lng = ~lng, lat =  ~lat, radius = ~circle_size,
                   popup = ~paste0(kommunenavn, 
                                   "<br>prosent_industri: ", round(prosent_industri,1)), 
                   color = "#ff1a1a", fillColor = "#ff1a1a", opacity = 0.8,
                   data = sf_datk %>% filter(prosent_industri >= 15))

```


